---
title: "Mantel Tests"
author: "Kyra Bankhead"
date: "2023-05-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Organize predictor and response variable dissimilarity matrices.

2. Run Mantel tests with different predictors.


```{r data, message=FALSE, warning=FALSE, include=FALSE}

# Set working directory here
setwd("C:/Users/bankh/My_Repos/Dolphins/data")

## load all necessary packages
require(ade4) # Look at Dai Shizuka/Jordi Bascompte
require(ncf) # For weights
require(vegan)

# Read file in to retain ILV
sample_data <- read.csv("sample_data.csv")
kov <- readRDS("kov.RDS")

# Read in social association matrix and data
nxn <- readRDS("nxn.RData")
list_years <- readRDS("list_years.RData")


```

## Transforming SRI similarity into distance

```{r dolpdist, echo=TRUE, message=FALSE, warning=FALSE}

# Transforming SRI similarity into distance
year <- 5
dolp_dist = nxn[[year]] + 0.00001
dolp_dist <- 1-nxn[[year]]
## Remove the redundant cells and the diagonal 
dolp_dist <- as.dist(dolp_dist)

```

## Organize data into foraging and HI frequency

```{r freq, echo=TRUE, message=FALSE, warning=FALSE}

# Select variables from the raw data
data <- list_years[[year]]
aux <- data[, c('Code', 'Behaviors', 'HumanInteraction', 'ConfHI')]

# Use 'Behaviors' variable to extract "Feed" and create another variable with two classes (Feed, Other)
aux$Foraging <- "Other"
aux$Foraging[grepl(pattern = 'Feed',
                   x = aux$Behaviors,
                   ignore.case = FALSE, perl = FALSE,
                   fixed = FALSE, useBytes = FALSE)] = "Feed"
#aux <- subset(aux, aux$Foraging == "Feed")
aux$ConfHI <- ifelse(aux$ConfHI == "0", 0, 1)

# Categorize ID to Foraging
IDbehav <- table(aux$Code, aux$Foraging)
IDbehav <- as.data.frame(IDbehav, stringsAsFactors = FALSE)
IDbehav <- IDbehav[,c(1,3)]
colnames(IDbehav) <- c("Code", "Forg_Freq")
# Group by the 'Code' column and sum the frequencies
IDbehav <- aggregate(. ~ Code, data = IDbehav, sum)

# HI behaviors should be partitioned into 3 different types
#' B = Begging (direct provisioning): F, G, H
#' P = patrolling/scavenging (indirect): A, B, C
#' D = foraging around fixed gear (humans not present):D, E, P
# Fix the code using ifelse statements
aux$ConfHI <- ifelse(aux$ConfHI %in% c("F", "G", "H"), "B",
                     ifelse(aux$ConfHI %in% c("A", "B", "C"), "P", 
                            ifelse(aux$ConfHI %in% c("D", "E", "P"), "D", "0")))

# Categorize ConfHI to IDs
rawHI <- as.matrix(table(aux$Code, aux$ConfHI))
rawHI <- as.data.frame(rawHI, stringsAsFactors = FALSE)
colnames(rawHI) <- c("Code", "ConfHI", "Freq")

## Add up the # of times each ID was seen in HI
HI <- "B"
IDbehav$HI <- rawHI$Freq[rawHI$ConfHI == HI & rawHI$ConfHI != "0"]
IDdata <- IDbehav
colnames(IDdata) <- c("Code", "Foraging", "HI")

## Proportion of time Foraging spent in HI
IDdata$HIprop <- as.numeric(IDdata$HI)/as.numeric(IDdata$Foraging)
IDdata[is.na(IDdata)] <- 0

# Only ID to prop
HIprop_ID <- IDdata[,c(1, 4)]

```

## Create HI Dissimularity Matrix using Eucladean Distance

```{r HIdis, echo=TRUE, message=FALSE, warning=FALSE}

# Dissimilarity of HI proportion among individual dolphins, using Euclidean distance
fake_HIprop <- HIprop_ID$HIprop
dissimilarity_HI <- as.matrix(dist(as.matrix(fake_HIprop), method = "euclidean"))
dissimilarity_HI[is.na(dissimilarity_HI)] <- 0

```

## Calculate Mantel test

```{r mantel, echo=TRUE, message=FALSE, warning=FALSE}

dissimilarity_HI <- as.dist(dissimilarity_HI) # HI dissimilarity
kov <- as.dist(kov) # Home range overlap

# Dissimilarity matrices
## HRO
hro_test <- mantel.rtest(dolp_dist, kov, nrepet = 1000)

plot(hro_test)

## HI
HI_test <- mantel.rtest(dolp_dist, dissimilarity_HI, nrepet = 1000)

plot(HI_test)

```

So far there are no correlation with HI engagement and associations.
