---
title: "Network Structure"
author: "Kyra Bankhead"
date: "2023-03-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Create the network structure from the association matrix.

2. Evaluate local and global network metrics.

3. Permutate the link weights using the WalkTrap algorithm.

4. Evaluate modularity.


# PART 1: *Network Structure*

```{r network, echo=TRUE, message=FALSE, warning=FALSE}
## load all necessary packages
require(igraph) # Look at Dai Shizuka/Jordi Bascompte
require(tnet) # For weights
require(sna)
require(statnet)

# Read in social association matrix
setwd("C:/Users/bankh/My_Repos/Dolphins/data")
# Read in social association matrix
nxn <- readRDS("nxn.RData")

# Test one year at a time
year <- 1
nxn <- nxn[[year]]

## Create social network
ig <- graph_from_adjacency_matrix(as.matrix(nxn),
                                  mode = c("undirected"),
                                  weighted = TRUE,
                                  diag = F, # No loops
                                  add.colnames = T,
                                  add.rownames = NA)

# Plot network
plot(ig,
     layout = layout_with_fr(ig),
     # link weight, rescaled for better visualization
     edge.width= E(ig)$weight*4,
     # node size as degree (rescaled)
     vertex.size= sqrt(igraph::strength(ig, vids = V(ig), mode = c("all"), loops = TRUE) *10 ),
     vertex.frame.color= NA, #"black",
     vertex.label.family = "Helvetica",
     vertex.label.color="black", 
     vertex.label.cex=0.8, 
     vertex.label.dist=2, 
     # edge.curved=0,
     vertex.frame.width=0.01,
)


```


# PART 2: *Network Metrics*

## Local Network Metrics

- Local clustering coefficient: Measure of the prevalence of node clusters in a network.

- Betweeness: A high betweenness means that the individual is in the communication path of other individuals, therefore, the individuals it interacts with, depend on its presence.

- Closeness: The larger the closeness centrality is for an individual, the more rapidly and easily it can influence the behavior of others.

- Degree: # Individual’s associates

- Strength: Total strength of an individuals’ associations

```{r lmetrics, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# Source edgelist function
source("../code/functions.R")

# Edgelist: Nodes (i & j) and edge (or link) weight
el <- matrix_to_edgelist(nxn, rawdata = FALSE, idnodes = FALSE)

# Centrality measures
# Weighted clustering coefficients
clustering_local_w(el, measure=c("am", "gm", "mi", "ma", "bi"))

## Betweenness centrality
betweenness_w(el, alpha=1)       

# Closeness centrality
closeness_w(el, alpha=1)

# Degree and strength centrality
degree_w(el, measure=c("degree","output"), type="out", alpha=1)


```

# Global Network Metrics

- Size: Number of nodes.

- Density/Connectance: Proportion of realized links (observed/possible links).

- Average Path Length (geodesic): Measures the shortest distance between two random nodes then average shortest pathways between all pairs of nodes. Shows how far apart any pair of individuals will be on average.

- Geodesic path: the shortest path through the network from one node to another (l).

- Diameter: Length of the longest geodesic path (d).

- Clustering coefficient: Tendency of nodes to cluster in the network (Are the friends’ friends also friends?).


```{r gmetrics, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#' Size
N = nrow(nxn)
#' Number of possible links: 
#' Nodes*(Nodes-1)/2: (-1 removes the node itself; /2 removes repetitions)
total = N*(N-1)/2
# Number of realized links: all non-zero cells in the association matrix
real = length(which(as.dist(nxn)!=0))
# Connectance: realized/total
real/total

# Shortest path lengths (geodesics) and diameter
# all binary shortest path lengths between nodes
distances(ig)

# mean shortest path
mean_distance(ig)

# Binary shortest path lengths
distance_table(ig, directed=FALSE)

# All Weighted Shortest path lenghts (or geodesics)
distance_w(el)
# weighted diameter
max(as.dist(distance_w(el)))

# Clustering coefficients
clustering_w (el, measure=c("am", "gm", "mi", "ma", "bi")) 

```


# PART 3: *Permutate Link Weights*

## Walktrap algorithm breakdown with one interation

```{r data, message=FALSE, warning=FALSE, include=FALSE}
## Edgelist
# Source edgelist function
source("../code/functions.R")
el <- matrix_to_edgelist(nxn, rawdata = FALSE, idnodes = FALSE)
## igraph format with weight
dolphin_ig <- graph.adjacency(as.matrix(nxn),mode="undirected",weighted=TRUE,diag=FALSE)
E(dolphin_ig)$weight

```


```{r oneit, message=FALSE, warning=FALSE, include=FALSE}
# Modularity by the WalkTrap algorithm 
dolphin_walk <- cluster_walktrap(dolphin_ig, weights = E(dolphin_ig)$weight, 
                                 steps = 4, merges = TRUE, modularity = TRUE, membership = TRUE)
## Modularity Q-value
modularity(dolphin_walk)
## Number of modules
groups(dolphin_walk)
## Membership of modules
membership(dolphin_walk)
## Save the edgelist into a new object
auxrand <- as.data.frame(el)
## Link weight distribution
auxrand$vw

# Permutate the link weights
sample(auxrand$vw)
## Save in the auxrand object
auxrand[,3] <- sample(auxrand$vw)

# Calculate the modularity Q-value for a new permutated edge list
## Create a network from the list of nodes
igrand <- graph.edgelist(el[,1:2]) 
### Add link weights
E(igrand)$weight <- el[,3]
### Make undirected graph
igrand <- as.undirected(igrand)
## Permutate the link weights
E(igrand)$weight <- sample(E(igrand)$weight)
## Calculate modularity Q-value
rmod <- walktrap.community(igrand)
modularity(rmod)
## Number of modules
groups(rmod)
## Membership of modules
membership(rmod)

# Difference from our empirical data?
modularity(dolphin_walk)
modularity(rmod)

```


## Permutate with multiple interations

```{r mult, echo=TRUE, message=FALSE, warning=FALSE}
# Run modularity permutations 1000 times
iter = 1000
randmod = numeric()
for(i in 1:iter){
  # Save the edgelist into a new object
  auxrand <- el
  # igraph format
  igrand <- graph.edgelist(auxrand[,1:2]) # Create a network from the list of nodes
  E(igrand)$weight <- auxrand[,3] # Add link weights
  igrand <- as.undirected(igrand) # Make undirected graph
  # Permutate the link weights
  E(igrand)$weight <- sample(E(igrand)$weight)
  # calculate the modularity Q-value
  rand_walk <- walktrap.community(igrand)
  randmod[i] <- modularity(rand_walk) # Save Q-value into a vector
}


## Calculate the 95% confidence interval (two-tailed test)
ci = quantile(randmod, probs=c(0.025, 0.975), type=2)

## Compare with the empirical Q-value
data.frame(Q=modularity(dolphin_walk), LowCI=ci[1], HighCI=ci[2])

## Visualization random Q distribution
hist(randmod, xlim=c(0,1))
### Empirical Q-value
abline(v= modularity(dolphin_walk), col="red")
### 2.5% CI
abline(v= ci[1], col="blue")
### 97.5% CI
abline(v= ci[2], col="blue")

```

We can reject the null hypothesis that individuals cluster at random and conclude that there is evidence that modularity is higher than what we would expect by chance. 


# PART 4: *Modularity*

- Newman's Q modularity: Stopping parameter Q removes links according to the betweenness. 

```{r modular, echo=TRUE, message=FALSE, warning=FALSE}
# Create a network from the first two columns
dolp_ig <- graph.edgelist(el[,1:2])
# Add the edge weights to this network by assigning an edge attribute called 'weight'.
E(dolp_ig)$weight <- as.numeric(el[,3])
# Create undirect network
dolp_ig <- as.undirected(dolp_ig)

# Plot
plot(dolp_ig, edge.width=E(dolp_ig)$weight*4, vertex.size=10, vertex.label=NA, edge.curved=F)

# Newman's Q modularity
newman <- cluster_leading_eigen(dolp_ig, steps = -1, weights = E(dolp_ig)$weight, 
                                start = NULL, options = arpack_defaults, callback = NULL, 
                                extra = NULL, env = parent.frame())

# Assign a random color to individuals of each module ('modules')
# Random color scheme
col <- rgb(runif(14), runif(14), runif(14))

V(dolp_ig)$color <- NA
for (i in 1:max(newman$membership)){
  sample(col)
  V(dolp_ig)$color[which(newman$membership==i)] = col[i]
}

plot(dolp_ig)

```

Since these modules can represent functional units, I need to test which mechanisms drive the modular topology by creating null models.