---
title: "Network Structure"
author: "Kyra Bankhead"
date: "2023-03-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Create the network structure from the association matrix.

2. Evaluate local and global network metrics.

3. Permutate the link weights using the WalkTrap algorithm.

4. Evaluate and plot modularity.


# PART 1: *Network Structure*

```{r data, message=FALSE, warning=FALSE, include=FALSE}
## load all necessary packages
require(igraph) # Look at Dai Shizuka/Jordi Bascompte
require(tnet) # For weights
require(sna)
require(statnet)
library(doParallel) # For faster computing

# Read in social association matrix
setwd("C:/Users/bankh/My_Repos/Dolphins/data")

# Add helpful functions
source("../code/functions.R") # edgelist function

# Read in social association matrix
nxn <- readRDS("../data/nxn.RData")
list_years <- readRDS("../data/list_years.RData")

```


```{r network, echo=TRUE, message=FALSE, warning=FALSE}

## Create social network
ig <- lapply(nxn, function (df) {
  graph_from_adjacency_matrix(
  df,
  mode = "undirected",
  weighted = TRUE,
  diag = FALSE)})

# Set the node names based on row names
row_names <- lapply(nxn, function (df) {rownames(df)})

for (i in seq_along(ig)) {
  V(ig[[i]])$name <- row_names[[i]]
}

## Only show IDs of HI dolphins
### subset_HI in "GLMM.R"
HI_data <-  diff_raw(subset_HI(list_years))
row_names_HI <- lapply(HI_data, function (df) {
  as.vector(df$Code[(df$DiffHI == "BG" | df$DiffHI == "SD" | 
                                    df$DiffHI == "FG") & df$Freq > 0])})

# Plot network
# Set up the plotting area with 1 row and 2 columns for side-by-side plots
par(mfrow=c(1, 2))

# Loop through the list of graphs and plot them side by side
for (i in 1:length(ig)) {
  plot(ig[[i]],
       layout = layout_with_fr(ig[[i]]),
       edge.width = E(ig[[i]])$weight * 4,
       vertex.size = sqrt(igraph::strength(ig[[i]], vids = V(ig[[i]]), mode = c("all"), loops = TRUE) * 10),
       vertex.frame.color = NA,
       vertex.label.family = "Helvetica",
       vertex.label = ifelse(V(ig[[i]])$name %in% row_names_HI[[i]], V(ig[[i]])$name, NA),
       vertex.label.color = "black",
       vertex.label.cex = 0.8,
       vertex.label.dist = 2,
       vertex.frame.width = 0.01)
}

# Reset the plotting area to its default configuration
par(mfrow=c(1, 1))

```


# PART 2: *Network Metrics*

## Local Network Metrics

- Local clustering coefficient: Measure of the prevalence of node clusters in a network (i.e. if dolphin associates all know each other, the clustering coefficient will be high).

- Betweeness: A high betweenness means that the individual is in the communication path of other individuals, therefore, the individuals it interacts with, depend on its presence.

- Closeness: The larger the closeness centrality is for an individual, the more rapidly and easily it can influence the behavior of others.

- Degree: # Individual’s associates

- Strength: Total strength of an individuals’ associations

```{r lmetrics, echo=TRUE, message=FALSE, warning=FALSE}

# Edgelist: Nodes (i & j) and edge (or link) weight
el <- readRDS("../data/el_years.RData")

# Set the node names based on row names
get_names <- function (matrix, metric) {
  row_names <- lapply(matrix, function (df) {rownames(df)})
for (i in seq_along(metric)) {
  metric[[i]][,1] <- row_names[[i]]
}
  return(metric)
  }

# Weighted clustering coefficients
cluster <- readRDS("../data/cluster.RData")
cluster_diffs <- get_names(nxn, cluster)
cluster_diffs_HI <- lapply(seq_along(cluster_diffs), function(i) {
  df <- cluster_diffs[[i]]
  df_new <- as.data.frame(df[df[, 1] %in% row_names_HI[[i]], , drop = FALSE])
  return(df_new)
})
compare_cluster <- merge(
  cluster_diffs_HI[[1]][, c(1, 2)], 
  cluster_diffs_HI[[2]][, c(1, 2)], 
  by.x = "node", 
  by.y = "node"
)
colnames(compare_cluster) <- c("ID", "Period.1", "Period.2")
compare_cluster[, c(2, 3)] <- sapply(compare_cluster[, c(2, 3)], as.numeric)
# Calculate differences
compare_cluster$Difference <- compare_cluster$Period.2 - compare_cluster$Period.1


# Betweenness centrality
between <- lapply(el, function (df) {betweenness_w(df, alpha=1)})
between_diffs <- get_names(nxn, between)
between_diffs_HI <- lapply(seq_along(between_diffs), function(i) {
  df <- between_diffs[[i]]
  df_new <- as.data.frame(df[df[, 1] %in% row_names_HI[[i]], , drop = FALSE])
  return(df_new)
})
compare_between <- merge(
  between_diffs_HI[[1]], 
  between_diffs_HI[[2]], 
  by.x = "node", 
  by.y = "node"
)
colnames(compare_between) <- c("ID", "Period.1", "Period.2")
compare_between[, c(2, 3)] <- sapply(compare_between[, c(2, 3)], as.numeric)
# Calculate differences
compare_between$Difference <- compare_between$Period.2 - compare_between$Period.1


# Closeness centrality
close <- lapply(el, function (df) {closeness_w(df, alpha=1)})
close_diffs <- get_names(nxn, close)
close_diffs_HI <- lapply(seq_along(close_diffs), function(i) {
  df <- close_diffs[[i]]
  df_new <- as.data.frame(df[df[, 1] %in% row_names_HI[[i]], , drop = FALSE])
  return(df_new)
})
compare_close <- merge(
  close_diffs_HI[[1]][, c(1, 2)], 
  close_diffs_HI[[2]][, c(1, 2)], 
  by.x = "node", 
  by.y = "node"
)
colnames(compare_close) <- c("ID", "Period.1", "Period.2")
compare_close[, c(2, 3)] <- sapply(compare_close[, c(2, 3)], as.numeric)
# Calculate differences
compare_close$Difference <- compare_close$Period.2 - compare_close$Period.1


# Degree and strength centrality
strength <- lapply(el, function (df) {degree_w(df, measure=c("degree","output"), type="out", alpha=1)})
strength_diffs <- get_names(nxn, strength)
strength_diffs_HI <- lapply(seq_along(strength_diffs), function(i) {
  df <- strength_diffs[[i]]
  df_new <- as.data.frame(df[df[, 1] %in% row_names_HI[[i]], , drop = FALSE])
  return(df_new)
})
compare_strength <- merge(
  strength_diffs_HI[[1]], 
  strength_diffs_HI[[2]], 
  by.x = "node", 
  by.y = "node"
)
colnames(compare_strength) <- c("ID", "Period.1_degree", "Period.1_strength", "Period.2_degree", "Period.2_strength")
compare_strength[, c(2:5)] <- sapply(compare_strength[, c(2:5)], as.numeric)
# Calculate differences
compare_strength$Difference_degree <- compare_strength$Period.2_degree - compare_strength$Period.1_degree
compare_strength$Difference_strength <- compare_strength$Period.2_strength - compare_strength$Period.1_strength

# Look at all of the local metrics together
## Add a column containing HI type
names_BG <- unlist(lapply(HI_data, function (df) {
  as.vector(df$Code[df$DiffHI == "BG" & df$Freq > 0])}))
names_SD <- unlist(lapply(HI_data, function (df) {
  as.vector(df$Code[df$DiffHI == "SD" & df$Freq > 0])}))
names_FG <- unlist(lapply(HI_data, function (df) {
  as.vector(df$Code[df$DiffHI == "FG" & df$Freq > 0])}))

HI_type <- ifelse(compare_cluster$ID %in% names_BG, "BG", 
                                   ifelse(compare_cluster$ID %in% names_SD, "SD", 
                                          ifelse(compare_cluster$ID %in% names_FG, "FG", "NA")))

local_metrics_HI <- as.data.frame(cbind(ID = compare_cluster$ID, HI_type = HI_type, Cluster_diff = compare_cluster$Difference, 
                               Between_diff = compare_between$Difference, Close_diff = compare_close$Difference, 
                               Degree_diff = compare_strength$Difference_degree, Strength_diff = compare_strength$Difference_strength))

## Add a rown to compare the averages of each metric with HI IDs
avg_metrics <- data.frame(ID = "Average", HI_type = "NA",
  Cluster_diff = mean(cluster[[2]][, 2]) - mean(cluster[[1]][, 2]),
  Between_diff = mean(between[[2]][, 2]) - mean(between[[1]][, 2]),
  Close_diff = mean(close[[2]][, 2]) - mean(close[[1]][, 2]),
  Degree_diff = mean(strength[[2]][, 2]) - mean(strength[[1]][, 2]),
  Strength_diff = mean(strength[[2]][, 3]) - mean(strength[[1]][, 3]))

local_metrics_HI <- merge(local_metrics_HI, avg_metrics, all = T)

# Make sure the metrics are numeric and HI_type is a factor
columns_to_convert <- c(3:7)  # Columns to be converted to numeric
local_metrics_HI[, columns_to_convert] <- sapply(local_metrics_HI[, columns_to_convert], as.numeric)
local_metrics_HI$HI_type <- as.factor(local_metrics_HI$HI_type)


```


# Plot different Metrics within HI types

```{r plotmetric, echo=TRUE, message=FALSE, warning=FALSE}

# Set up the plotting area with 1 row and 5 columns
par(mfrow = c(1, 5))

# Get different metric values in a vector
metric_value <- colnames(local_metrics_HI[, columns_to_convert])

# Define the desired order for HI types
desired_order <- c("NA", "BG", "FG", "SD")

# Loop over each metric
for (i in seq_along(columns_to_convert)) {
  
  # Extract the metric for the current column
  metric <- local_metrics_HI[, columns_to_convert[i]]
  
  # Convert HI_type to factor with the desired order
  local_metrics_HI$HI_type <- factor(local_metrics_HI$HI_type, levels = desired_order)
  
  # Group the metric by HI_type and order the names
  grouped_metric <- split(metric, local_metrics_HI$HI_type)
  
  # Set up colors based on HI type
  colors <- rainbow(length(names(grouped_metric)))
  
  # Create the boxplot
  boxplot(
    grouped_metric,
    names = names(grouped_metric),
    col = colors,
    notch = FALSE,
    main = metric_value[i],
    border = "black"
  )
  # Draw a horizontal line at y = 0
  abline(h = 0, col = "red")
}


```



# Global Network Metrics

- Size: Number of nodes.

- Density/Connectance: Proportion of realized links (observed/possible links).

- Average Path Length (geodesic): Measures the shortest distance between two random nodes then average shortest pathways between all pairs of nodes. Shows how far apart any pair of individuals will be on average.

- Geodesic path: the shortest path through the network from one node to another (l).

- Diameter: Length of the longest geodesic path (d).

- Clustering coefficient: Tendency of nodes to cluster in the network (Are the friends’ friends also friends?).


```{r gmetrics, echo=TRUE, message=FALSE, warning=FALSE}

#' Breakdown: connectance = length(which(as.dist(orca_hwi)!=0))/(N*(N-1)/2)
## Calculate connectance for each matrix
calculate_connectance <- function(matrix) {
  N <- nrow(matrix)
  total <- N * (N - 1) / 2
  real <- sum(matrix != 0)  # Count non-zero elements
  connectance <- real / total
  return(connectance)
}

connectance_list <- lapply(nxn, calculate_connectance)
connectance_list

# Shortest path lengths (geodesics) and diameter
# # mean shortest path
dist <- lapply(ig, function(df) {mean_distance(df)})
dist

```


# PART 3: *Permutate Link Weights*

## Walktrap algorithm breakdown with one interation

```{r edgedata, message=FALSE, warning=FALSE, include=FALSE}
## Edgelist
## igraph format with weight
n.cores <- detectCores()
system.time({
  registerDoParallel(n.cores)
  dolphin_ig <- list()
  for (j in seq_along(list_years)) {
    dolphin_ig[[j]] <- graph.adjacency(as.matrix(nxn[[j]]),
                                       mode="undirected",
                                       weighted=TRUE, diag=FALSE)
  }  
  ### End parallel processing
  stopImplicitCluster()
})

```


```{r oneit, message=FALSE, warning=FALSE, include=FALSE}
# Modularity by the WalkTrap algorithm 
system.time({
  registerDoParallel(n.cores)
  dolphin_walk <- list()
  for (k in seq_along(list_years)) {
    dolphin_walk[[k]] <- cluster_walktrap(dolphin_ig[[k]], weights = E(dolphin_ig[[k]])$weight, 
                                     steps = 4, merges = TRUE, modularity = TRUE, membership = TRUE)
  } 
  ### End parallel processing
  stopImplicitCluster()
})


```


## Permutate with multiple interations

```{r mult, echo=TRUE, message=FALSE, warning=FALSE}
# Run modularity permutations 1000 times for each matrix
run_mod <- function(el_list, dolphin_walk_list) {
  iter <- 1000
  randmod <- numeric(iter)  # Initialize a numeric vector to store Q-values
  
  for (i in 1:iter) {
    # Save the edgelist into a new object and permutate the link weights
    auxrand <- el_list
    auxrand[, 2] <- sample(auxrand[, 2])
    
    # Create an igraph graph from the permuted edgelist
    igrand <- graph.edgelist(as.matrix(auxrand[, 1:2]), directed = FALSE)
    E(igrand)$weight <- auxrand[, 2]  # Assign link weights
    
    # Calculate modularity using walktrap community detection
    rand_walk <- walktrap.community(igrand)
    randmod[i] <- modularity(rand_walk)  # Save Q-value into the vector
  }
  
  # Calculate the 95% confidence interval (two-tailed test)
  ci <- quantile(randmod, probs = c(0.025, 0.975), type = 2)
  
  # Visualization of the random Q distribution
  hist(randmod, xlim = c(0, 0.6), main = "Random Q Distribution", xlab = "Q-value", ylab = "Frequency", col = "lightblue")
  
  # Empirical Q-value
  abline(v = modularity(dolphin_walk_list), col = "red")
  
  # 2.5% CI
  abline(v = ci[1], col = "blue")
  
  # 97.5% CI
  abline(v = ci[2], col = "blue")
  
  # Return a data frame with Q-value and confidence intervals
  result <- data.frame(Q = modularity(dolphin_walk_list), LowCI = ci[1], HighCI = ci[2])
  return(result)
}

# Plot null distributions for modularity
# Set up the plotting area with 2 rows and 2 column
par(mfrow=c(2, 1))
run_mod(el_list = el[[1]], dolphin_walk_list = dolphin_walk[[1]])
run_mod(el_list = el[[2]], dolphin_walk_list = dolphin_walk[[2]])


```

We can reject the null hypothesis that individuals cluster at random and conclude that there is evidence that modularity is higher than what we would expect by chance. 


# PART 4: *Modularity*

- Newman's Q modularity: Stopping parameter Q removes links according to the betweenness. 

```{r modular, echo=TRUE, message=FALSE, warning=FALSE}
# Create an unweighted network
system.time({
  registerDoParallel(n.cores)
  dolp_ig <- list()
  for (l in seq_along(list_years)) {
    dolp_ig[[l]] <- graph.edgelist(el[[l]][,1:2])
    # Add the edge weights to this network
    E(dolp_ig[[l]])$weight <- as.numeric(el[[l]][,3])
    # Create undirected network
    dolp_ig[[l]] <- as.undirected(dolp_ig[[l]])
  }   
  ### End parallel processing
  stopImplicitCluster()
})

# Newman's Q modularity
newman <- lapply(dolp_ig, function (df) {cluster_leading_eigen(df, steps = -1, weights = E(df)$weight, 
                                start = NULL, options = arpack_defaults, callback = NULL, 
                                extra = NULL, env = parent.frame())})


# Set the node names based on row names
BG <- SD <- FG <- vector("list", length = length(dolp_ig))

for (i in seq_along(dolp_ig)) {
  # Set the node names
  V(dolp_ig[[i]])$name <- rownames(nxn[[i]])
  
  ## Parse out what HI behavior they engage in
  BG[[i]] <- as.vector(HI_data[[i]]$Code[HI_data[[i]]$DiffHI == "BG" & HI_data[[i]]$Freq > 0])
  SD[[i]] <- as.vector(HI_data[[i]]$Code[HI_data[[i]]$DiffHI == "SD" & HI_data[[i]]$Freq > 0])
  FG[[i]] <- as.vector(HI_data[[i]]$Code[HI_data[[i]]$DiffHI == "FG" & HI_data[[i]]$Freq > 0])
  
  ## Initialize label_color attribute for each node
  V(dolp_ig[[i]])$label_color <- "black"
  
  ## Make a different text color for each category
  V(dolp_ig[[i]])$label_color[V(dolp_ig[[i]])$name %in% BG[[i]]] <- "red"  
  V(dolp_ig[[i]])$label_color[V(dolp_ig[[i]])$name %in% SD[[i]]] <- "yellow" 
  V(dolp_ig[[i]])$label_color[V(dolp_ig[[i]])$name %in% FG[[i]]] <- "blue"
}

# Generate a vector of colors based on the number of unique memberships
for (i in seq_along(dolp_ig)) {
  V(dolp_ig[[i]])$color <- NA
  col <- rainbow(max(newman[[i]]$membership))
  
  for (j in 1:max(newman[[i]]$membership)){
    V(dolp_ig[[i]])$color[which(newman[[i]]$membership==j)] <- col[j]
  }
}

# Make sure the HI dolphins stand out
for (i in seq_along(dolp_ig)) {
  V(dolp_ig[[i]])$size <- ifelse(V(dolp_ig[[i]])$name %in% row_names_HI[[i]], 10, 5)
}

# Set up the plotting area with 1 row and 2 columns for side-by-side plots
par(mfrow=c(1, 2))
# Main labels for the plots
main_labels <- c("1993-2004 Network", "2005-2014 Network")  # Replace with appropriate main labels

# Plot the graph with individual IDs as labels
for (i in seq_along(dolp_ig)) {
plot(dolp_ig[[i]],
     layout = layout_with_fr(dolp_ig[[i]]),
     # link weight, rescaled for better visualization
     edge.width= E(dolp_ig[[i]])$weight*4,
     # node size as degree (rescaled)
     vertex.size= V(dolp_ig[[i]])$size,
     vertex.frame.color= NA, #"black",
     vertex.label.family = "Helvetica",
     vertex.label=ifelse(V(dolp_ig[[i]])$name %in% row_names_HI[[i]], as.character(V(dolp_ig[[i]])$name), NA), 
     vertex.label.color = V(dolp_ig[[i]])$label_color, 
     vertex.label.cex=0.8, 
     vertex.label.dist=0.5, 
     # edge.curved=0,
     vertex.frame.width=0.01)
  # Add the main label above the plot
  title(main = main_labels[i], line = -1)
  }

```

Since these modules can represent functional units, I need to test which mechanisms drive the modular topology by creating null models.